Dynamics of Laterally Propagating Flames in X-ray Bursts. I. Burning Front Structure

### 2019-11-26 xrb/paper.tex
add sponge

When we begin the simulation, there is a transient phase as the flame
gets established.  Material that is forced upward will encounter the
steep density gradient at the top of the atmosphere and accelerate as
it is blown out of the atmosphere.  Eventually this material will fall
back to the top of the atmosphere.  In this paper, we are mostly
concerned with the behavior of the flame and not any material that is
violently blown out of the top of the atmosphere, so we apply a sponge
to this region.  This is similar to the method we previously used in \cite{xrb3d}.
The sponge drives the velocity of the material in the low
density \MarginPar{add mathematical form of sponge} regions at the top
of the atmosphere to zero.  This sponging helps increase our timestep
as well.

>>>>

When we begin the simulation, there is a transient phase as the flame
gets established.  Material that is forced upward will encounter the
steep density gradient at the top of the atmosphere and accelerate as
it is blown out of the atmosphere.  Eventually this material will fall
back to the top of the atmosphere.  In this paper, we are mostly
concerned with the behavior of the flame and not any material that is
violently blown out of the top of the atmosphere, so we apply a sponge
to this region.  This is similar to the method we previously used in \cite{xrb2}, and takes
the form of a source term to the momentum and energy equations of the form:
\begin{align}
{\bf S}_{\rho \Ub} &= \rho \Ub \frac{f}{\Delta t} \\
S_{\rho E} &= \rho \Ub \cdot {\bf S}_{\rho \Ub}
\end{align}
with the sponge forcing $f$ dependent on the density.  We define the sponge
shape as:
\begin{equation}
s = \left \{
     \begin{array}{cc}
            0   & \rho > \rho_\mathrm{upper} \\
            \frac{1}{2}
                \left [ 1 - \cos \left ( \frac{\pi (\rho - \rho_\mathrm{upper})}{\Delta \rho} \right ) \right ]  & \rho_\mathrm{upper} \ge \rho > \rho_\mathrm{lower} \\
                     1 & \rho < \rho_\mathrm{lower} 
             \end{array} \right .
\end{equation}
Here $\rho_\mathrm{upper}$ and $\rho_\mathrm{lower}$ are the densities where the sponge transitions to being fully applied.  We take $\rho_\mathrm{upper} = 10^2~\gcc$ and $\rho_\mathrm{lower} = 1~\gcc$, with $\Delta \rho = \rho_\mathrm{upper} - \rho_\mathrm{lower}$.
The sponge update is done implicitly to get the effective forcing, $f$:
\begin{equation}
f = -\left [ 1 - \frac{1}{1 + \alpha s} \right ]
\end{equation}
with $\alpha = \Delta t/\tau_\mathrm{sponge}$.  Here
$\tau_\mathrm{sponge}$ is the timescale over which the sponge acts.
We take $\tau_\mathrm{sponge} = 10^{-7}~\mathrm{s}$.
The sponge drives the velocity of the material in the low
density regions at the top
of the atmosphere to zero.  This sponging helps increase our timestep
as well.
